---
title: "Traffic and Pedestrian Data"
author: "Tiasia Saunders"
date: "2026-06/01"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---


- Pitch Story Memo (plan to update 06/02/2026 -- spent most of time on Monday figuring out methodology for cacluation of the veil of darkness)

- Initial Pitch: Pitch 1: Is Racial Bias in Traffic Stops a Chicago Problem or a Statewide Problem? 
	Using four years of Illinois traffic stop data from 2021 to 2024, I want to examine whether racial disparities in traffic stops extend beyond the Chicago Police Department to suburban and downstate Illinois law enforcement agencies. While CPD has faced significant scrutiny, especially since the fatal shooting of Dexter Reed, a 26-year-old Black man who was shot and killed in 2024 by the Chicago Police Department during a shootout between him and officers, smaller departments across Illinois operate with far less oversight and have received almost no data-driven accountability reporting. 
- By examining this through the veil of darkness lens, which controls for visibility conditions to isolate racial bias in stop decisions, across multiple Illinois agencies, I can determine whether this is a Chicago-specific failure or a statewide structural problem. If suburban departments, county sheriffs, and state police show the same pattern as CPD, that fundamentally reframes the conversation around racial bias in Illinois policing and raises urgent questions about what oversight, if any, exists for the departments that fly under the radar. 
- While racial disparities in policing have been documented nationally, most public attention in Illinois focuses on the Chicago Police Department. If similar patterns appear across suburban, county, and state law enforcement agencies, it would suggest that racial bias is not isolated to one department but may be embedded throughout Illinois policing. This matters because many of these agencies receive far less public scrutiny, media coverage, and oversight despite conducting thousands of traffic stops each year.


- Veil of Darkness Explanation: 

- It is a statistical methodology, created by Jeffrey Grogger and Greg Ridgeway in 2006, that is used to test for racial bias in traffic stops
- It compares daytime stops to nighttime stops, relying on the hypothesis that at nighttime, natural darkness masks a driver’s race, which prevents racially biased officers from profiling them 
- The theory: if racial profiling exists, there should be a higher percentage of minority drivers stopped during daylight hours when officers can easily identify race, compared to nighttime hours when the ‘veil of darkness’ conceals the person’s race 
- 5 PM–9 PM only (most academic papers use this time range since I can measure stops taken during daylight, during sunset, and after sunset across the entire year) 
- Late-March - Mid-September 
Nighttime: 9 pm 
- Late September - Mid-March 
Nighttime: 5 pm 

- Proposed Methodology: (How will I get the evidence needed to prove or disprove my hypothesis) 
Benchmark: Still working on how to do this, but I need to figure out how to determine a ‘benchmark’ of how to evaluate whether the reason behind mostly black people getting stopped may be due to the high percentage of black people living in that said area ( I think to mitigate this, I will examine the data through the veil of darkness angle to determine potential racial bias) 
- Veil of Darkness: Will be determining racial bias through looking at the two sets of time ranges, the day and the night, to examine whether there is an obvious decrease in the amount of Black or brown people who are stopped at night due to the lack of visibility 
Suspicion Threshold: Will be looking at the resulting actions of stops after the initial traffic stop to determine whether there is a drastic disparity between how Black/brown versus white people are treated during a traffic stop 
- Will be looking at whether police officers have sufficient or suspected reason to search for evidence, and whether the ‘hit rate’ (that the officers do find something illegal) for minorities is relatively lower than their white counterparts or around the same, to determine if the threshold is equal for both 
Pretexual Lens: Will be looking at whether minority races are stopped for low-level administrative reasons, such as broken taillight, or expired registration, which are discretionary reasons for stops and can be largely pretextual, in that the officer uses these reasons to search for other (worse) crimes. 

- Data Process Workflow: 
 - So, I have downloaded the 16 (four per year from 2021-2024) datasets into my .rmd file and have begun preliminary analysis on it to see what I am working with 
 - I have added the 2020 Census  (let me know if there is a better source that could be used for this data) racial population makeup to my RMD.file to be able to calculate the per capita rates to determine which race is truly most impacted by traffic stops 
  - I plan to determine the racial discrimination through two factors: 
  - The veil of darkness angle 
  -  Examining the breakdown of the driver’s race, reason for the stop, and the result of the stop to determine if there are disparities there 

- Hypothesis: 
- I suspect that I will find a pattern within the broader Illinois as a whole, that most of the law enforcement agencies (including CPD) tend to stop Black or minority race people more often in the daytime than at night, due to being able to identify the race in the daytime. Additionally, I suspect that there is a racial disparity between the initial reasoning for the traffic stop and the potential search and overall conclusion of the stop (citation, written warning, or verbal warning) between minority and white drivers. 

- Potential Problems with the Hypothesis/ Core Arguments: 

- Quick (two sentences) overview of what the story is about:
- If racial disparities exist across Illinois agencies regardless of location or department size, it would suggest that concerns about biased policing extend far beyond Chicago and may require broader statewide accountability measures.

- The Why: 
- If racial disparities exist across Illinois agencies regardless of location or department size, it would suggest that concerns about biased policing extend far beyond Chicago and may require broader statewide accountability measures.
- Target Audience:
  - Illinois residents and drivers
  - Communities disproportionately affected by traffic enforcement
  - Police Department officials/ law enforcement officers 
  - Policymakers and criminal justice reform advocates
  - Local government officials and police oversight bodies

- Three Potential Interviewees: 
  - A person impacted by racial discrimination during a traffic stop 
  - Plan to look on social media for stories on TikTok, Reddit, or Instagram to be able to find an individual for this 
  - A representative from the Illinois Department of Transportation or Illinois State Police familiar with statewide traffic stop reporting requirements.
  - A civil rights attorney, policing researcher, or advocate who studies racial disparities in traffic enforcement.

- Data Records: 
  - I already received the data from the Investigative Project on Race and Equity, and it includes data from 2021-2024. Each year has four datasets, two from the Chicago Police Department data (traffic and pedestrian data) and two from other Illinois police departments' data (traffic and pedestrian data. 
  - I will be bringing in the 2020 Census Illinois population racial breakdown data to be able to calculate the per capita rates to be able to have a true reflection of who police traffic stops impact the most. 

- Visual Ideas: 
  - Statewide map showing racial disparities in stop rates by agency.
  - Bar chart comparing racial composition of daytime versus nighttime stops.
  - Scatterplot comparing stop disparities and search hit rates by department.
  - An interactive dashboard allowing readers to compare agencies.
  - Timeline showing trends in racial disparities from 2021 to 2024

- Impact: 
  - This project could reveal whether racial disparities in traffic stops are a problem across Illinois, not just within the Chicago Police Department. The findings could help inform conversations about police accountability and bring attention to departments that conduct thousands of traffic stops each year but rarely face public scrutiny.


- Resources: 
  - https://news.stanford.edu/stories/2020/05/veil-darkness-reduces-racial-bias-traffic-stops
  - https://5harad.com/papers/100M-stops.pdf	







```{r include=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(googlesheets4)
library(googledrive)
library(janitor)
library(rio)
library(data.table)
library(googledrive)
library(knitr)
```

# READ CAREFULLY

# Import Data: READ CAREFULLY

# READ CAREFULLY

Data is coming from a shared Google Drive folder due to the large file
sizes There is a two-step process: PART ONE: First, download the data to
YOUR HARD DRIVE. YOU RUN THIS SEQUENCE ONCE. PART TWO: Load into R. This
is how you will load the data for future analysis.

There are two files from the Chicago Police Department -- "CPD" are the
.xlsx files -- and two from surrounding police departments, the .txt
files

One of these files -- traffic_CPD_21 -- is too large to push to GitHub
unless you use the lfs (large file size) protocol. It is copied below if
you want to try it. I don't recommend it. It's time consuming and I have
provided an acceptable workaround.

I would just keep the large data files on your GitHub repo folder on
your hard drive. Don't try to push them to GitHub

#### Data Tables Definitions

-   traffic_x: data of traffic stops happening under other agencies,
    other than Chicago police department
-   traffic_CPD_x: data of traffic stops happening under Chicago police
    department
-   pedestrian_redacted_x: data of pedestrian stops happening under
    other agencies, other than Chicago police department
-   pedestrian_CPD_x: data of pedestrian stops happening under Chicago
    police department
    
Date Column: year-month-date format 

# PART ONE - download the data to YOUR HARD DRIVE. YOU RUN THIS SEQUENCE ONCE.

### 2021 Data 

```{r}
#Enter '1' to start a new auth process or select a pre-authorized account.
#1: Send me to the browser for a new auth process.
#CPD - 2021 REDCATED Traffic Data
#https://docs.google.com/spreadsheets/d/1xFOnp3Dx11JzH6h1UAPIMU8RVThFvIDh/edit?usp=sharing&ouid=102324743793755798467&rtpof=true&sd=true
googledrive::drive_auth()
drive_download(
  as_id("1xFOnp3Dx11JzH6h1UAPIMU8RVThFvIDh"),  # from the share link
  path = "../data/traffic_CPD_21.xlsx",
  overwrite = TRUE
)
## commenting the below line out because it is too big to pass through to github
##traffic_CPD_21 <- read_excel("../data/traffic_CPD_21.xlsx")|> 
  ##clean_names()

#CPD - 2021 REDACTED Pedestrian Data
#https://docs.google.com/spreadsheets/d/1GrmIDziDBGn2e1fiXLKiDQCcZLIe6S90/edit?usp=sharing&ouid=102324743793755798467&rtpof=true&sd=true
drive_download(
  as_id("1GrmIDziDBGn2e1fiXLKiDQCcZLIe6S90"),  # from the share link
  path = "../data/pedestrian_CPD_21.xlsx",
  overwrite = TRUE
)
pedestrian_CPD_21 <- read_excel("../data/pedestrian_CPD_21.xlsx")|> 
  clean_names()

#2021 Traffic Data Redacted
#https://drive.google.com/file/d/1q5D_MGMProIshXZV_h2RlIiZUsYU14hp/view?usp=sharing
drive_download(
  as_id("1q5D_MGMProIshXZV_h2RlIiZUsYU14hp"),  # from the share link
  path = "../data/traffic_21.txt",
  overwrite = TRUE
)
traffic_21 <- fread("../data/traffic_21.txt", sep = "~")|> 
  clean_names()


#2021 Pedestrian Data Redacted
#https://drive.google.com/file/d/1sXRtyPLyE2HCKfZi2zQxb_z-fe0KFday/view?usp=sharing

drive_download(
  as_id("1sXRtyPLyE2HCKfZi2zQxb_z-fe0KFday"),  # from the share link
  path = "../data/pedestrian_redacted_21.txt",
  overwrite = TRUE
)
pedestrian_redacted_21 <- fread("../data/pedestrian_redacted_21.txt", sep = "~")|> 
  clean_names()


```


### 2022 Data 

```{r eval=FALSE, include=TRUE}
### load the 2022- 2024 data 

drive_download(
  as_id("1GQAcNcBpyDpeH9Ge9IYB0lnefI637NAl"),  # from the share link
  path = "../data/traffic_CPD_22.xlsx",
  overwrite = TRUE
)
traffic_CPD_22 <- read_excel("../data/traffic_CPD_22.xlsx")|> 
  clean_names()


drive_download(
  as_id("17beRQ5as-mECg-EgmUe999caWEDInMDI"),  # from the share link
  path = "../data/pedestrian_CPD_22.xlsx",
  overwrite = TRUE
)
pedestrian_CPD_22 <- read_excel("../data/pedestrian_CPD_22.xlsx")|> 
  clean_names()


# https://drive.google.com/file/d/17MCVZSBX67VKy-HwIVUOShA8_2gY7cHZ/view?usp=drive_link
drive_download(
  as_id(""),  # from the share link
  path = "../data/traffic_22_pt1.txt",
  overwrite = TRUE
)
traffic_22_pt1 <- fread("../data/traffic_22_pt1.txt", sep = "~")|> 
  clean_names()

drive_download( 
  as_id("1xU-lzfk6k9yDfVsp7CzKSze_17kYpA04"),  # from the share link
  path = "../data/traffic_22_pt2.txt",
  overwrite = TRUE
)
traffic_22_pt2 <- fread("../data/traffic_22_pt2.txt", sep = "~")|> 
  clean_names()

drive_download(
  as_id("1VTSg-4_RSOG8VVhpBBsrhWWzxcLIMXUB"),  # from the share link
  path = "../data/pedestrian_redacted_22.txt",
  overwrite = TRUE
)
pedestrian_redacted_22 <- fread("../data/pedestrian_redacted_22.txt", sep = "~")|> 
  clean_names()


```

### 2023 Data

```{r eval=FALSE, include=TRUE}

drive_download(
  as_id("10yCnj0-AY86q2nSgctO7-GhT1sZhUU_y"),  # from the share link
  path = "../data/traffic_CPD_23.xlsx",
  overwrite = TRUE
)
traffic_CPD_23 <- read_excel("../data/traffic_CPD_23.xlsx")|> 
  clean_names()

drive_download(
  as_id("1w6vnVrrQ3IzXMbl3wdk-tUsaHt2vWgfA"),
  path = "/Users/tiasiasaunders/Desktop/code/tiasia-saunders/data/pedestrian_CPD_23.xlsx",
  overwrite = TRUE
)

pedestrian_CPD_23 <- read_excel("../data/pedestrian_CPD_23.xlsx") |> 
  clean_names()

drive_download(
  as_id("1MWRhVQLD7652Ln9TiJEOBphEVIabd4kJ"),  # from the share link
  path = "../data/traffic_23.txt",
  overwrite = TRUE
)
traffic_23 <- fread("../data/traffic_23.txt", sep = "~")|> 
  clean_names()

drive_download(
  as_id("1qZaCB_YD0Q31RvmfyeVzjyP0aGqP_xZD"),  # from the share link
  path = "../data/pedestrian_redacted_23.txt",
  overwrite = TRUE
)
pedestrian_redacted_23 <- fread("../data/pedestrian_redacted_23.txt", sep = "~")|> 
  clean_names()

```

### 2024 Data

```{r eval=FALSE, include=TRUE}


drive_download(
  as_id("1pqA4y9-Upgol5xWxfDvift43j24ZJZzI"),  # from the share link
  path = "../data/traffic_CPD_24.xlsx",
  overwrite = TRUE
)
traffic_CPD_24 <- read_excel("../data/traffic_CPD_24.xlsx")|> 
  clean_names()

drive_download(
  as_id("1cqwJTHZnae5NbLdgQhVUNttHFnCkugMC"),  # from the share link
  path = "../data/pedestrian_CPD_24.xlsx",
  overwrite = TRUE
)
pedestrian_CPD_24 <- read_excel("../data/pedestrian_CPD_24.xlsx")|> 
  clean_names()

drive_download(
  as_id("1N-GdlJqGQXsRqjW6xF8zTjVvdNuWu0g1"),  # from the share link
  path = "../data/traffic_24.txt",
  overwrite = TRUE
)
traffic_24 <- fread("../data/traffic_24.txt", sep = "~")|> 
  clean_names()


drive_download(
  as_id("1nSGllC3zosxWIrh-8W_wvrPa-NWtvoOV"),  # from the share link
  path = "../data/pedestrian_redacted_24.txt",
  overwrite = TRUE
)
pedestrian_redacted_24 <- fread("../data/pedestrian_redacted_24.txt", sep = "~")|> 
  clean_names()


```

```{r}
setwd("/Users/tiasiasaunders/Desktop/code/tiasia-saunders/data")
```

```{r}
## setting working directory to root of project so I can pull in the data files 
knitr::opts_knit$set(root.dir = "/Users/tiasiasaunders/Desktop/code/tiasia-saunders")
```



# PART TWO: Load into R. This is how you will load the data for future analysis.

```{r eval=FALSE, include=TRUE}

traffic_CPD_21 <- read_excel("../data/traffic_CPD_21.xlsx")|> 
  clean_names()
pedestrian_CPD_21 <- read_excel("../data/pedestrian_CPD_21.xlsx")|> 
  clean_names()
traffic_21 <- fread("../data/traffic_21.txt", sep = "~")|> 
  clean_names()
pedestrian_redacted_21 <- fread("../data/pedestrian_redacted_21.txt", sep = "~")|> 
  clean_names()
traffic_CPD_22 <- read_excel("../data/traffic_CPD_22.xlsx") |> 
  clean_names()
pedestrian_CPD_22 <- read_excel("../data/pedestrian_CPD_22.xlsx") |> 
  clean_names()
traffic_22_pt1 <- fread("../data/traffic_22_pt1.txt") |> 
  clean_names ()
traffic_22_pt2 <- fread("../data/traffic_22_pt2.txt") |>
  clean_names()
pedestrian_redacted_22 <- fread("../data/pedestrian_redacted_22.txt") |> 
  clean_names() 
traffic_CPD_23 <- read_excel("../data/traffic_CPD_23.xlsx") |>
  clean_names()
pedestrian_CPD_23 <- read_excel("../data/pedestrian_CPD_23.xlsx") |>
  clean_names() 
traffic_23 <- fread("../data/traffic_23.txt") |>
  clean_names()
pedestrian_redacted_23 <- fread("../data/pedestrian_redacted_23.txt") |>
  clean_names()
traffic_CPD_24 <- read_excel("../data/traffic_CPD_24.xlsx") |>
  clean_names()
pedestrian_CPD_24 <- read_excel("../data/pedestrian_CPD_24.xlsx") |> 
  clean_names() 
traffic_24 <- fread("../data/traffic_24.txt") |>
  clean_names()
pedestrian_redacted_24 <- fread("../data/pedestrian_redacted_24.txt") |> 
  clean_names()

```


### Preliminary Basic Analysis 

```{r}
pedestrian_CPD_21 |> 
  group_by(datestop) |> 
  count() |> 
  arrange(desc(n))

```

```{r}
traffic_21 |> 
  count(agency_name, agency_code) |> 
  arrange(desc(n))

```
- There are 786 other law enforcement agencies in the traffic_2021 data 


```{r}
## creating a df with agency_name and agency_code do I can add agency names to my pt 1 and pt 2 of traffic_2022 data 
agency_lookup <- traffic_21 |> 
  select(agency_name, agency_code) |>
  distinct() 
```

```{r eval=FALSE, include=TRUE}
## commenting this out since I cant pass it through github 
## combining pt 1 and 2 of traffic_2022 data so it can be uniformed with the other datasets 
## traffic_22 <- bind_rows(
##  traffic_22_pt1,
 ## traffic_22_pt2
##)
```

```{r eval=FALSE, include=TRUE}
traffic_22 <- traffic_22 |> 
  left_join(
    agency_lookup |> select(agency_code, agency_name),
    by = "agency_code"
  )

```


```{r eval=FALSE, include=TRUE}
## ensuring join worked correctly 
traffic_22 |> 
  select(agency_code, agency_name)

```


```{r eval=FALSE, include=TRUE}
traffic_22 |> 
  count(agency_name, agency_code) |> 
  arrange(desc(n))
```

- There are 792 other law enforcement agencies in the traffic_2022 data 

```{r eval=FALSE, include=TRUE}
traffic_23 |> 
  count(agency_name) |> 
  arrange(desc(n))
```

- There are 804 other law enforcement agencies in the traffic_2021 data 

```{r eval=FALSE, include=TRUE}
traffic_24 |> 
  count(agency_name) |> 
  arrange(desc(n))
```

- There are 786 other law enforcement agencies in the traffic_2024 data 

Testing this part of my hypothesis: Additionally, I suspect that there is a racial disparity between the initial reasoning for the traffic stop and the potential search and overall conclusion of the stop (citation, written warning, or verbal warning) between minority and white drivers. 

```{r eval=FALSE, include=TRUE}
## getting breakdown of reason for stop by race 
traffic_21 |>
  count(driver_race, reason_for_stop,result_of_stop, sort = TRUE)
```


```{r eval=FALSE, include=TRUE}
traffic_21 |>
  count(driver_race, reason_for_stop) |>
  group_by(driver_race) |>
  mutate(percent = n / sum(n) * 100) |>
  arrange(driver_race, desc(percent))

traffic_21 |>
  count(driver_race, result_of_stop) |>
  group_by(driver_race) |>
  mutate(percent = n / sum(n) * 100) |>
  arrange(driver_race, desc(percent))
```

### So right now it seems like white people get stopped more, but we have to account for per_capita rates due to the population demographic in Illnois. So, I will be bringing in 2020 Illinois Racial Breakdown Census Data to perform the per_capita caculations. 
- Because Illinois's racial composition changed relatively little between 2020 and 2024, 2020 U.S. Census population counts were used as a common denominator for all years analyzed. Using a single population benchmark allows differences in traffic stop rates across years to reflect changes in enforcement patterns rather than fluctuations in annual population estimates.



- Plan for 06/02/26: Bringing in 2020 Census Illnois Racial Makeup of Population to be able to get the per capita rates for traffic stops 






```{r eval=FALSE, include=TRUE}
pedestrian_CPD_21 |> 
  select(pedsex, pedrace, search_bey_pat, search_bey, search_bey_by) |> 
  count(pedrace) |> 
   group_by(pedrace) |> 
  arrange(desc(n))
```

- So, in 2021 there were 48,044 Black pedestrain stops and 5,009 white pedestrain stops, and 14,611 Hispanic stops. 















# Pokin around (Robs part)

```{r}
CDP_ped_stops <- pedestrian_CPD_21|> 
  group_by(datestop) |> 
  count() |> 
  arrange(desc(n))

head(CDP_ped_stops)
#Most stops were on Oct 9, 364 stops

```

#average stops

```{r}
CDP_ped_stops |> 
  ungroup() |> 
  summarize(mean(n))

#Average day has 188 stops 
```

# Number of days above / below average

```{r}
CDP_ped_stops |> 
  mutate(Context = case_when(
         n > 187 ~ "more",
         n < 187 ~ "less")) |> 
  group_by(Context) |> 
  count()

#186 days had fewer than the average 187 stops and 177 days had more
```

```{r}
traffic_21 |> 
  group_by(agency_name) |> 
  count() |> 
  arrange(desc(n))

#ILLINOIS STATE POLICE had the most traffic stops with	223643	
```




