---
title: "Traffic and Pedestrian Data"
author: "Tiasia Saunders"
date: "2026-06/01 UPDATE THE DATE"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

### Chicago Police Department versus other Illinois Law Enforcement Agencies Correlation with the Veil of Darkness

**Need a lead image or graphic.**

**emphasize your main findings on top and then go into the explanation
of the project**

- Pitch: Is Racial Bias in Traffic Stops a Chicago Problem or a
  Statewide Problem?

  - Using four years of Illinois traffic stop data from 2021 to 2024, I
    want to examine whether racial disparities in traffic stops extend
    beyond the Chicago Police Department to suburban and downstate
    Illinois law enforcement agencies. While CPD **write Chicago Police
    Dept. instead of CPD throughout** has faced significant scrutiny,
    especially since the fatal shooting of Dexter Reed, a 26-year-old
    Black man who was shot and killed in 2024 by the Chicago Police
    Department during a shootout between him and officers, smaller
    departments across Illinois operate with far less oversight and have
    received almost no data-driven accountability reporting.
  - By examining this through the veil of darkness lens, which controls
    for visibility conditions to isolate racial bias in stop decisions,
    across multiple Illinois agencies, I can determine whether this is a
    Chicago-specific failure or a statewide structural problem. If
    suburban departments, county sheriffs, and state police show the
    same pattern as CPD, that fundamentally re-frames the conversation
    around racial bias in Illinois policing and raises urgent questions
    about what oversight, if any, exists for the departments that fly
    under the radar.
  - While racial disparities in policing have been documented
    nationally, most public attention in Illinois focuses on the Chicago
    Police Department. If similar patterns appear across suburban,
    county, and state law enforcement agencies, it would suggest that
    racial bias is not isolated to one department but may be embedded
    throughout Illinois policing. This matters because many of these
    agencies receive far less public scrutiny, media coverage, and
    oversight despite conducting thousands of traffic stops each year.

- Veil of Darkness Explanation:

  - It is a statistical methodology, created by Jeffrey Grogger and Greg
    Ridgeway in 2006, that is used to test for racial bias in traffic
    stops **rsw comment: link to their study**
  - It compares daytime stops to nighttime stops, relying on the
    hypothesis that at nighttime, natural darkness masks a driver’s
    race, which prevents racially biased officers from profiling them
  - The theory: if racial profiling exists, there should be a higher
    percentage of minority drivers stopped during daylight hours when
    officers can easily identify race, compared to nighttime hours when
    the ‘veil of darkness’ conceals the person’s race
  - **rsw comment: Need to explain that these times are just estimates
    you are using for a proof of concept and that you will nail down the
    precise sunrise, sunset times using National Weather Service
    historical data later**
  - Daytime: 6 a.m. to 5 p.m.
  - Nighttime: 10 p.m. to 4 a.m.
  - Excluded transition hours: 4 a.m.–5:59 a.m. and 5 p.m.–9:59 p.m.

- Proposed Methodology: (How will I get the evidence needed to prove or
  disprove my hypothesis)

  - Will be looking at the racial breakdown of traffic stops across all
    Illinois law enforcement agencies, including CPD, to determine if
    there is a racial disparity in who gets stopped the most.
  - So analysis will be done on ~~traffic_24 (~~all other Illinois law
    enforcement agencies traffic enforcement data) and t~~raffic_CPD_24
    (~~Chicago Police Department traffic enforcement data) to determine
    if there is a racial disparity in both CPD and the other Illinois
    law enforcement agencies, and then comparing the two to see if there
    is a similar pattern in both data sets.

- **RSW COMMENT: YOU ALREAD SAID THIS:** ~~Veil of Darkness: Will be
  determining racial bias through looking at the two sets of time
  ranges, the day and the night, to examine whether there is an obvious
  decrease in the amount of Black or brown people who are stopped at
  night due to the lack of visibility Suspicion Threshold: Will be
  looking at the resulting actions of stops after the initial traffic
  stop to determine whether there is a drastic disparity between how
  Black/brown versus white people are treated during a traffic stop~~

  - Will be looking at whether police officers have sufficient or
    suspected reason to search for evidence, and whether the ‘hit rate’
    (that the officers do find something illegal) for minorities is
    relatively lower than their white counterparts or around the same,
    to determine if the threshold is equal for both

  - Will be examining racial disparities in the initial reason for the
    stop, and then the resulting actions of the stop (citation, written
    warning, verbal warning) to determine if there is a disparity there
    as well

- Data Process Workflow: **rsw comment: i tabbed these bullet points in
  so you have a heading and sub points. this wasn't organized properly -
  it was all subheadings**

  - ~~So,~~ I have downloaded the 16 (four per year from 2021-2024)
    datasets into my .rmd file and have begun preliminary analysis on it
    to see what I am working with

  - ~~For the sake of time, I will be performing the analysis~~ **To
    start, I analyze** the 2024 traffic data. Might, look at pedestrian
    due to curiosity if time permits.

  - I have added the 2020 Census **(get the 2024 ACS Census data)**
    l~~et me know if there is a better source that could be used for
    this data)~~ racial population makeup to my RMD.file to be able to
    calculate the per capita rates to determine which race is truly most
    impacted by traffic stops

    **rsw comment: There's a lot of repetition in this narrative.
    Consolidate this and be more careful about your writing. It seems
    rushed and unorganized**

  - ~~I plan to determine the racial discrimination through three
    factors:~~

  - ~~The veil of darkness angle~~

  - ~~Examining the breakdown of the driver’s race, reason for the stop,
    and the result of the stop to determine if there are disparities
    there~~

  - ~~Examining the reason for stop and hit rates (whether officers find
    something illegal during a search) for minorities versus white
    people to determine if there is a disparity in the suspicion
    threshold for minorities versus white people)~~

- Hypothesis:

  - I suspect that I will find a pattern within the broader Illinois as
    a whole, that most of the law enforcement agencies (including CPD)
    tend to stop Black or minority race people more often in the daytime
    than at night, due to being able to identify the race in the
    daytime. Additionally, I suspect that there is a racial disparity
    between the initial reasoning for the traffic stop and the potential
    search and overall conclusion of the stop (citation, written
    warning, or verbal warning) between minority and white drivers.

- Potential Problems with the Hypothesis/ Core Arguments:

- Quick (two sentences) overview of what the story is about:

- If racial disparities exist across Illinois agencies regardless of
  location or department size, it would suggest that concerns about
  biased policing extend far beyond Chicago and may require broader
  statewide accountability measures.

- The Why:

  - If racial disparities exist across Illinois agencies regardless of
    location or department size, it would suggest that concerns about
    biased policing extend far beyond Chicago and may require broader
    statewide accountability measures.

- Target Audience:

  - Illinois residents and drivers
  - Communities disproportionately affected by traffic enforcement
  - Police Department officials/ law enforcement officers
  - Policymakers and criminal justice reform advocates
  - Local government officials and police oversight bodies

- Three Potential Interviewees:

  - A person impacted by racial discrimination during a traffic stop
  - Plan to look on social media for stories on TikTok, Reddit, or
    Instagram to be able to find an individual for this
  - A representative from the Illinois Department of Transportation or
    Illinois State Police familiar with statewide traffic stop reporting
    requirements.
  - A civil rights attorney, policing researcher, or advocate who
    studies racial disparities in traffic enforcement.

- Data Records:

  - I already received the data from the Investigative Project on Race
    and Equity, and it includes data from 2021-2024. Each year has four
    datasets, two from the Chicago Police Department data (traffic and
    pedestrian data) and two from other Illinois police departments'
    data (traffic and pedestrian data.
  - I will be bringing in the 2020 Census Illinois population racial
    breakdown data to be able to calculate the per capita rates to be
    able to have a true reflection of who police traffic stops impact
    the most.

- Visual Ideas:

  - Statewide map showing racial disparities in stop rates by agency.
  - Bar chart comparing racial composition of daytime versus nighttime
    stops.
  - Scatterplot comparing stop disparities and search hit rates by
    department.
  - An interactive dashboard allowing readers to compare agencies.
  - Timeline showing trends in racial disparities from 2021 to 2024

- Impact:

  - This project could reveal whether racial disparities in traffic
    stops are a problem across Illinois, not just within the Chicago
    Police Department. The findings could help inform conversations
    about police accountability and bring attention to departments that
    conduct thousands of traffic stops each year but rarely face public
    scrutiny.

- Resources:

  - <https://news.stanford.edu/stories/2020/05/veil-darkness-reduces-racial-bias-traffic-stops>
  - <https://5harad.com/papers/100M-stops.pdf>

#### Loading Libraries

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/tiasiasaunders/Desktop/code/tiasia-saunders")
```

```{r}
library(tidyverse)
library(here)
library(readxl)
library(googlesheets4)
library(googledrive)
library(janitor)
library(rio)
library(data.table)
library(googledrive)
library(knitr)
library(ggplot2)
library(flourishcharts)
library(scales)
library(dplyr)
library(tidyr)
library(DT) 
library(dplyr)
library(ggrepel)
library(lubridate)
library(readxl)
library(tidytext)

```

##### Data Tables Definitions

- traffic_x: data of traffic stops happening under other agencies, other
  than Chicago police department
- traffic_CPD_x: data of traffic stops happening under Chicago police
  department
- pedestrian_redacted_x: data of pedestrian stops happening under other
  agencies, other than Chicago police department
- pedestrian_CPD_x: data of pedestrian stops happening under Chicago
  police department

##### Data that I am using for this analysis:

- traffic_2024: data of other law enforcement agencies happening under
  different agencies in Illinois
- traffic_CPD_2024: data of traffic stops happening under Chicago police
  department

```{r eval=FALSE, include=FALSE}
##### 2021 Data 
#Enter '1' to start a new auth process or select a pre-authorized account.
#1: Send me to the browser for a new auth process.
#CPD - 2021 REDCATED Traffic Data
#https://docs.google.com/spreadsheets/d/1xFOnp3Dx11JzH6h1UAPIMU8RVThFvIDh/edit?usp=sharing&ouid=102324743793755798467&rtpof=true&sd=true
googledrive::drive_auth()
drive_download(
  as_id("1xFOnp3Dx11JzH6h1UAPIMU8RVThFvIDh"),  # from the share link
  path = "../data/traffic_CPD_21.xlsx",
  overwrite = TRUE
)
## commenting the below line out because it is too big to pass through to github
##traffic_CPD_21 <- read_excel("../data/traffic_CPD_21.xlsx")|> 
  ##clean_names()

#CPD - 2021 REDACTED Pedestrian Data
#https://docs.google.com/spreadsheets/d/1GrmIDziDBGn2e1fiXLKiDQCcZLIe6S90/edit?usp=sharing&ouid=102324743793755798467&rtpof=true&sd=true
drive_download(
  as_id("1GrmIDziDBGn2e1fiXLKiDQCcZLIe6S90"),  # from the share link
  path = "../data/pedestrian_CPD_21.xlsx",
  overwrite = TRUE
)
pedestrian_CPD_21 <- read_excel("../data/pedestrian_CPD_21.xlsx")|> 
  clean_names()

#2021 Traffic Data Redacted
#https://drive.google.com/file/d/1q5D_MGMProIshXZV_h2RlIiZUsYU14hp/view?usp=sharing
drive_download(
  as_id("1q5D_MGMProIshXZV_h2RlIiZUsYU14hp"),  # from the share link
  path = "../data/traffic_21.txt",
  overwrite = TRUE
)
traffic_21 <- fread("../data/traffic_21.txt", sep = "~")|> 
  clean_names()


#2021 Pedestrian Data Redacted
#https://drive.google.com/file/d/1sXRtyPLyE2HCKfZi2zQxb_z-fe0KFday/view?usp=sharing

drive_download(
  as_id("1sXRtyPLyE2HCKfZi2zQxb_z-fe0KFday"),  # from the share link
  path = "../data/pedestrian_redacted_21.txt",
  overwrite = TRUE
)
pedestrian_redacted_21 <- fread("../data/pedestrian_redacted_21.txt", sep = "~")|> 
  clean_names()


```

```{r eval=FALSE, include=FALSE}
##### 2022 Data 

### load the 2022- 2024 data 

drive_download(
  as_id("1GQAcNcBpyDpeH9Ge9IYB0lnefI637NAl"),  # from the share link
  path = "../data/traffic_CPD_22.xlsx",
  overwrite = TRUE
)
traffic_CPD_22 <- read_excel("../data/traffic_CPD_22.xlsx")|> 
  clean_names()


drive_download(
  as_id("17beRQ5as-mECg-EgmUe999caWEDInMDI"),  # from the share link
  path = "../data/pedestrian_CPD_22.xlsx",
  overwrite = TRUE
)
pedestrian_CPD_22 <- read_excel("../data/pedestrian_CPD_22.xlsx")|> 
  clean_names()


# https://drive.google.com/file/d/17MCVZSBX67VKy-HwIVUOShA8_2gY7cHZ/view?usp=drive_link
drive_download(
  as_id(""),  # from the share link
  path = "../data/traffic_22_pt1.txt",
  overwrite = TRUE
)
traffic_22_pt1 <- fread("../data/traffic_22_pt1.txt", sep = "~")|> 
  clean_names()

drive_download( 
  as_id("1xU-lzfk6k9yDfVsp7CzKSze_17kYpA04"),  # from the share link
  path = "../data/traffic_22_pt2.txt",
  overwrite = TRUE
)
traffic_22_pt2 <- fread("../data/traffic_22_pt2.txt", sep = "~")|> 
  clean_names()

drive_download(
  as_id("1VTSg-4_RSOG8VVhpBBsrhWWzxcLIMXUB"),  # from the share link
  path = "../data/pedestrian_redacted_22.txt",
  overwrite = TRUE
)
pedestrian_redacted_22 <- fread("../data/pedestrian_redacted_22.txt", sep = "~")|> 
  clean_names()


```

```{r eval=FALSE, include=FALSE}
##### 2023 Data
drive_download(
  as_id("10yCnj0-AY86q2nSgctO7-GhT1sZhUU_y"),  # from the share link
  path = "../data/traffic_CPD_23.xlsx",
  overwrite = TRUE
)
traffic_CPD_23 <- read_excel("../data/traffic_CPD_23.xlsx")|> 
  clean_names()

drive_download(
  as_id("1w6vnVrrQ3IzXMbl3wdk-tUsaHt2vWgfA"),
  path = "/Users/tiasiasaunders/Desktop/code/tiasia-saunders/data/pedestrian_CPD_23.xlsx",
  overwrite = TRUE
)

pedestrian_CPD_23 <- read_excel("../data/pedestrian_CPD_23.xlsx") |> 
  clean_names()

drive_download(
  as_id("1MWRhVQLD7652Ln9TiJEOBphEVIabd4kJ"),  # from the share link
  path = "../data/traffic_23.txt",
  overwrite = TRUE
)
traffic_23 <- fread("../data/traffic_23.txt", sep = "~")|> 
  clean_names()

drive_download(
  as_id("1qZaCB_YD0Q31RvmfyeVzjyP0aGqP_xZD"),  # from the share link
  path = "../data/pedestrian_redacted_23.txt",
  overwrite = TRUE
)
pedestrian_redacted_23 <- fread("../data/pedestrian_redacted_23.txt", sep = "~")|> 
  clean_names()

```

#### 2024 Chicago + Law Enforcement Traffic Stop Data Analysis

- So, for the sake of my analysis, I am using 2024 Chicago Police
  Department and other Illinois law enforcement agencies. I am
  attempting to find racial disparities through the per capita of 100k
  people in total stops, reason for stop, and result of stop.
  Additionally, I am categorizing time frames (10 pm - 4 am = nighttime,
  6 am - 5 pm = daytime) to examine the disparities in the data through
  the veil of darkness angle.

```{r eval=FALSE, include=TRUE}
googledrive::drive_auth()
drive_download(
  as_id("1pqA4y9-Upgol5xWxfDvift43j24ZJZzI"),  # from the share link
  path = "../data/traffic_CPD_24.xlsx",
  overwrite = TRUE
)
traffic_CPD_24 <- read_excel("../data/traffic_CPD_24.xlsx")|> 
  clean_names()

drive_download(
  as_id("1cqwJTHZnae5NbLdgQhVUNttHFnCkugMC"),  # from the share link
  path = "../data/pedestrian_CPD_24.xlsx",
  overwrite = TRUE
)
pedestrian_CPD_24 <- read_excel("../data/pedestrian_CPD_24.xlsx")|> 
  clean_names()

drive_download(
  as_id("1N-GdlJqGQXsRqjW6xF8zTjVvdNuWu0g1"),  # from the share link
  path = "../data/traffic_24.txt",
  overwrite = TRUE
)
traffic_24 <- fread("../data/traffic_24.txt", sep = "~")|> 
  clean_names()

drive_download(
  as_id("1nSGllC3zosxWIrh-8W_wvrPa-NWtvoOV"),  # from the share link
  path = "../data/pedestrian_redacted_24.txt",
  overwrite = TRUE
)
pedestrian_redacted_24 <- fread("../data/pedestrian_redacted_24.txt", sep = "~")|> 
  clean_names()


```

##### Loading Datasets

```{r}
dir.create("data", showWarnings = FALSE)
```

```{r eval=FALSE, include=TRUE}
googledrive::drive_auth()

# ALWAYS create folder in working directory (not ../data)
dir.create("data", showWarnings = FALSE, recursive = TRUE)

drive_download(
  as_id("1pqA4y9-Upgol5xWxfDvift43j24ZJZzI"),
  path = file.path("data", "traffic_CPD_24.xlsx"),
  overwrite = TRUE
)

traffic_CPD_24 <- read_excel("data/traffic_CPD_24.xlsx") |> 
  clean_names()

drive_download(
  as_id("1N-GdlJqGQXsRqjW6xF8zTjVvdNuWu0g1"),
  path = file.path("data", "traffic_24.txt"),
  overwrite = TRUE
)

traffic_24 <- fread("data/traffic_24.txt", sep = "~") |> 
  clean_names()
```

```{r}
traffic_CPD_24 <- read_excel("data/traffic_CPD_24.xlsx") |> clean_names()
traffic_24 <- fread("data/traffic_24.txt", sep="~") |> clean_names()
```

```{r eval=FALSE, include=FALSE}
getwd()
```

```{r eval=FALSE, include=FALSE}

traffic_CPD_21 <- read_excel("../data/traffic_CPD_21.xlsx")|> 
  clean_names()
pedestrian_CPD_21 <- read_excel("../data/pedestrian_CPD_21.xlsx")|> 
  clean_names()
traffic_21 <- fread("../data/traffic_21.txt", sep = "~")|> 
  clean_names()
pedestrian_redacted_21 <- fread("../data/pedestrian_redacted_21.txt", sep = "~")|> 
  clean_names()
traffic_CPD_22 <- read_excel("../data/traffic_CPD_22.xlsx") |> 
  clean_names()
pedestrian_CPD_22 <- read_excel("../data/pedestrian_CPD_22.xlsx") |> 
  clean_names()
traffic_22_pt1 <- fread("../data/traffic_22_pt1.txt") |> 
  clean_names ()
traffic_22_pt2 <- fread("../data/traffic_22_pt2.txt") |>
  clean_names()
pedestrian_redacted_22 <- fread("../data/pedestrian_redacted_22.txt") |>   clean_names() 
traffic_CPD_23 <- read_excel("../data/traffic_CPD_23.xlsx") |>
  clean_names()
pedestrian_CPD_23 <- read_excel("../data/pedestrian_CPD_23.xlsx") |>
  clean_names() 
traffic_23 <- fread("../data/traffic_23.txt") |>
  clean_names()
pedestrian_redacted_23 <- fread("../data/pedestrian_redacted_23.txt") |>
  clean_names()
traffic_CPD_24 <- read_excel("../data/traffic_CPD_24.xlsx") |>
  clean_names()
pedestrian_CPD_24 <- read_excel("../data/pedestrian_CPD_24.xlsx") |> 
  clean_names() 
traffic_24 <- fread("../data/traffic_24.txt") |>
  clean_names()
pedestrian_redacted_24 <- fread("../data/pedestrian_redacted_24.txt") |>   clean_names()
census_2020_illinois <- read_excel("../data/census_2020_illinois_data.xlsx") |>
  clean_names()

```

```{r eval=FALSE, include=FALSE}
pedestrian_CPD_21 |> 
  group_by(datestop) |> 
  count() |> 
  arrange(desc(n))



traffic_21 |> 
  count(agency_name, agency_code) |> 
  arrange(desc(n))

# - There are 786 other law enforcement agencies in the traffic_2021 data 
```

```{r eval=FALSE, include=FALSE}
## creating a df with agency_name and agency_code do I can add agency names to my pt 1 and pt 2 of traffic_2022 data 
agency_lookup <- traffic_21 |> 
  select(agency_name, agency_code) |>
  distinct() 
```

```{r eval=FALSE, include=FALSE}
## commenting this out since I cant pass it through github 
## combining pt 1 and 2 of traffic_2022 data so it can be uniformed with the other datasets 
## traffic_22 <- bind_rows(
##  traffic_22_pt1,
 ## traffic_22_pt2
##)
```

```{r eval=FALSE, include=FALSE}
traffic_22 <- traffic_22 |> 
  left_join(
    agency_lookup |> select(agency_code, agency_name),
    by = "agency_code"
  )

```

```{r eval=FALSE, include=FALSE}
## ensuring join worked correctly 
traffic_22 |> 
  select(agency_code, agency_name)

```

```{r eval=FALSE, include=FALSE}
traffic_22 |> 
  count(agency_name, agency_code) |> 
  arrange(desc(n))
# - There are 792 other law enforcement agencies in the traffic_2022 data 
```

```{r eval=FALSE, include=FALSE}
traffic_23 |> 
  count(agency_name) |> 
  arrange(desc(n))
# - There are 804 other law enforcement agencies in the traffic_2021 data 
```

#### Preliminary Basic Analysis

- Redefining my columns by their associated meaning according to the
  traffic stop data dictionary, so I can have a better understanding of
  the data and be able to do more analysis on it.

```{r}
# re coding the race and gender numerical values into what they stand for according to the traffic stop data dictionary 

traffic_CPD_24 <- traffic_CPD_24 |>
  mutate(
    sex = recode(as.character(drsex),
      "1" = "Male",
      "2" = "Female",
      "3" = "Non-Binary", 
      "9" = "Unknown"
    ),
    race = recode(as.character(drrace),
      "1" = "White",
      "2" = "Black",
      "3" = "American Indian",
      "4" = "Hispanic",
      "5" = "Asian",
      "6" = "NHPI"
    ),
    vehicle_search = recode(as.character(vehsrchcond), 
      "1" = "Yes",  
      "2" = "No" 
    ), 
    vehicle_drugs_found = recode(as.character(vehdrugs),
      "0" = "Not applicable",
      "1" = "Yes",
      "2" = "No"
    ), 
    reason_for_stop = recode(as.character(reasstop), 
      "1" = "Moving Violation", 
      "2" = "Equipment", 
      "3" = "License Plate/ Registration",
      "4" = "Commercial Vehicle"
    ), 
    result_of_stop = recode(as.character(resstop),
      "1" = "Citation",
      "2" = "Written Warning", 
      "3" = "Verbal Warning (stop card)"
    )
  )

  
```

```{r}

# re coding the race and gender numerical values into what they stand for according to the traffic stop data dictionary 
# re coding the race and gender numerical values into what they stand for according to the traffic stop data dictionary 

traffic_24$driver_sex <- dplyr::recode(
  as.character(traffic_24[["driver_sex"]]),
  "1" = "Male",
  "2" = "Female",
  "3" = "Non-Binary",
  "9" = "Unknown"
)

traffic_24$driver_race <- dplyr::recode(
  as.character(traffic_24[["driver_race"]]),
  "1" = "White",
  "2" = "Black",
  "3" = "American Indian",
  "4" = "Hispanic",
  "5" = "Asian",
  "6" = "NHPI"
)

traffic_24$vehicle_search <- dplyr::recode(
  as.character(traffic_24[["vehicle_search_conducted"]]),
  "1" = "Yes",
  "2" = "No"
)

traffic_24$vehicle_drugs_found <- dplyr::recode(
  as.character(traffic_24[["vehicle_drugs_found"]]),
  "0" = "Not applicable",
  "1" = "Yes",
  "2" = "No"
)

traffic_24$reason_for_stop <- dplyr::recode(
  as.character(traffic_24[["reason_for_stop"]]),
  "1" = "Moving Violation",
  "2" = "Equipment",
  "3" = "License Plate/ Registration",
  "4" = "Commercial Vehicle"
)

traffic_24$result_of_stop <- dplyr::recode(
  as.character(traffic_24[["result_of_stop"]]),
  "1" = "Citation",
  "2" = "Written Warning",
  "3" = "Verbal Warning (stop card)"
)
```

```{r}
gender_CPD_24 <- traffic_CPD_24 |> 
  count(sex, race) |> 
  arrange(desc(n))

gender_CPD_24

race_CPD_24 <- traffic_CPD_24 |> 
  count(race) |> 
  arrange(desc(n))

race_CPD_24

gender_race_CPD_24 <- traffic_CPD_24 |> 
  count(sex, race) |> 
  arrange(desc(n))

gender_race_CPD_24
```

```{r}

gender_race_CPD_24 <- gender_race_CPD_24 |>
  arrange(desc(n)) |>
  mutate(
    highlight = if_else(row_number() <= 3, "Top 3", "Other"),
    group = paste(sex, race)
  )

ggplot(
  gender_race_CPD_24,
  aes(
    x = reorder(group, n),
    y = n,
    fill = highlight
  )
) +
  geom_col() +
  
  geom_text(
    aes(label = comma(n)),
    hjust = -0.1,
    size = 3
  ) +
  
  coord_flip() +
  
  scale_fill_manual(
    values = c(
      "Top 3" = "#B22222",
      "Other" = "gray70"
    )
  ) +
  
  scale_y_continuous(
    labels = comma,
    expand = expansion(mult = c(0, .15))
  ) +
  
  labs(
    title = "Chicago Police Department Traffic stops by race and sex",
    subtitle = "Top three race-gender groups highlighted",
    x = NULL,
    y = "Number of stops",
    caption = "Source: Chicago Police Department • Graphic by Tiasia Saunders"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold")
  )
```

- In 2024, there were 206,148 men who were involved in a CPD traffic
  stop, while 86,591 women were stopped, and 4ll non-binary people were
  stopped.
- As for race, in descending order Black people were stopped 130,938
  times, followed by Hispanic/ Latino with 101,891 stops, and then white
  people were stopped 43,304 times.
- When broken down by gender and race, it was found that the top three
  gender/racial demographics who were stopped by CPD in 2024 were Black
  men with 85,522 stops, Hispanic men with 76,917 stops, and Black women
  with 45,225 stops.

```{r}
## going to do code of datestop time and duration, then make graphic showing the timeline 
timeline_monthly <- traffic_CPD_24 |>
  mutate(month = floor_date(as.Date(datestop), "month")) |>
  count(month)

max_stops <- max(timeline_monthly$n)

ggplot(timeline_monthly, aes(x = month, y = n)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +

  geom_vline(
    xintercept = as.Date("2024-03-21"),
    linetype = "dashed",
    color = "red"
  ) +
  annotate(
    "label",
    x = as.Date("2024-03-21"),
    y = max_stops * 0.9,
    label = "Dexter Reed fatal\nshooting incident",
    size = 2
  ) +

  geom_vline(
    xintercept = as.Date("2024-08-21"),
    linetype = "dashed",
    color = "red"
  ) +
  annotate(
    "label",
    x = as.Date("2024-08-21"),
    y = max_stops * 0.75,
    label = "Injustice Watch investigation:\nNearly 200,000 undocumented arrests",
    size = 2
  ) +

  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b"
  ) +
  scale_y_continuous(
    labels = scales::comma
  ) +
  labs(
    title = "Chicago Police Department traffic stops by month, 2024",
    x = NULL,
    y = "Number of stops",
    caption = "Source: Chicago Police Department • Graphic by Tiasia Saunders"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(hjust = 1, size = 8, color = "gray40")
  )

```

- So, I looked into an explanation as to why there was a steady decrease
  of traffic stops by the CPD in 2024, so once looking it up I found
  this
  <a href="https://boltsmag.org/chicago-police-secret-traffic-stops/">article</a>,
  which explained that there were 210,622 undocumented traffic stops in
  2024, despite a 2003 law requiring them to document every traffic
  stop.
- In addition, the Dexter Reed incident in which a 26-year old Black man
  died at the result of a shootout incident between him and policemen in
  late March of 2024, may have resulted in heavier scrutiny due to
  advocacy groups and civil rights lawsuits.
  - The incident had stemmed from a seatbelt violation traffic stop.

```{r}
traffic_CPD_24 |> 
  count(city_i) |> 
  arrange(desc(n))

write.csv(
  traffic_CPD_24,
  "traffic_CPD_24.csv",
  row.names = FALSE
)
```

- The CPD made traffic stops in a little over 5,000 (5,083) locations in
  2024

```{r}
## looking at vehicle make and model, the ones most likely to be stopped 

vehicle_make_race <- traffic_CPD_24 |>
  count(vehmake, race) |>
  pivot_wider(
    names_from = race,
    values_from = n,
    values_fill = 0
  ) 


vehicle_make_race
```

**sort this datatable descending by the most common vehicle and just
display the first 10 entries**

```{r}
DT::datatable(
  vehicle_make_race,
  caption = "Traffic stops by vehicle make and driver race",
  filter = "top",
  rownames = FALSE,
  options = list(
    pageLength = 25,
    scrollX = TRUE
  )
)
```

- So, we could just look at the top three Chevrolet cars have been
  stopped for Black and Hispanic the most, followed by white (its almost
  10,000 less then their Black counterpart), and then Asian, which is
  only in the hundreds.
- The highest for white people is with the Toyota model at 6,685 stops,
  but even then black people were stopped more often in these cars at
  7,707 stops, and Hispanic were stopped the most with 10,584 stops.

```{r eval=FALSE, include=FALSE}
## looking at which demographic has reason for stop and then have to submit to drug search or dog search , ## looking at the hit rate of drug searches, and which race is more likely to have something found during a drug search

## holding off on this, will be examining it through the veil of darkness and the per capita caculations 
##traffic_CPD_24 |>
  ##count(race, reasonforstop, vehsrchcondby, vehdrugs)
```

#### Preliminary Data Analysis With the Other Illinois Law Enforcement Agencies for 2024

```{r eval=FALSE, include=TRUE}
traffic_24 |> 
  count(agency_name) |> 
  arrange(desc(n)) |> 
  distinct()
```

- There are 786 other law enforcement agencies in the traffic_2024 data.

```{r}
gender_traffic_24 <- traffic_24 |> 
  count(driver_sex, driver_race) |> 
  arrange(desc(n))

gender_traffic_24

race_traffic_24 <- traffic_24 |> 
  count(driver_race) |> 
  arrange(desc(n))

race_traffic_24

gender_race_traffic_24 <- traffic_24 |> 
  count(driver_sex, driver_race) |> 
  arrange(desc(n))

gender_race_traffic_24

```

**the headline is cut off on the ggplot**

\*\***i think the main thing is to normalize this graphic and the CPD
graphic by total demographic. we would expect a significant number of
black people to be pulled over in Chicago because the city is 28% Black.
out in the suburbs, probably few Black folks. let's try to address
that**

```{r}
## creating the bar chart 

gender_race_traffic_24 <- gender_race_traffic_24 |>
  arrange(desc(n)) |>
  mutate(
    highlight = if_else(row_number() <= 3, "Top 3", "Other"),
    group = paste(driver_sex, driver_race)
  )

ggplot(
  gender_race_traffic_24,
  aes(
    x = reorder(group, n),
    y = n,
    fill = highlight
  )
) +
  geom_col() +
  
  geom_text(
    aes(label = comma(n)),
    hjust = -0.1,
    size = 3
  ) +
  
  coord_flip() +
  
  scale_fill_manual(
    values = c(
      "Top 3" = "#B22222",
      "Other" = "gray70"
    )
  ) +
  
  scale_y_continuous(
    labels = comma,
    expand = expansion(mult = c(0, .15))
  ) +
  
  labs(
    title = "Illinois Police Departments (excluded CPD) Traffic stops by race and sex",
    subtitle = "Top three race-sex groups highlighted",
    x = NULL,
    y = "Number of stops",
    caption = "Source: FOIA Request Results  • Graphic by Tiasia Saunders"
  ) +
  
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold")
  )

```

- So, when I am looking at this bar chart of all the other police
  departments traffic stops, I notice the top two are white males with
  583,079 stops and white females with 371,651, and followed by Hispanic
  males with 236,865 stops. At first glance, it looks like white people
  are being stopped more than their minority counterparts, but as I
  hypothesized, I believe this is due to the population size of white
  people being very big in the state of Illinois. So, I am now going to
  account for per capital calculations with both the CPD to ensure that
  minorities are the most likely to be stopped and see if it lies true
  for the other local law enforcement agencies.

```{r}
## creating the timeline with annotations 

timeline_traffic_monthly <- traffic_24 |>
  mutate(
    date_of_stop = as.Date(.data[["date_of_stop"]]),
    month = lubridate::floor_date(date_of_stop, unit = "month")
  ) |>
  count(month, name = "n") |>
  arrange(month)

max_stops <- max(timeline_traffic_monthly$n, na.rm = TRUE)

ggplot(timeline_traffic_monthly, aes(x = month, y = n)) +
  
  geom_line(linewidth = 1) +
  geom_point(size = 2) +

  geom_vline(
    xintercept = as.Date("2024-03-21"),
    linetype = "dashed",
    color = "red"
  ) +

  annotate(
    "label",
    x = as.Date("2024-03-21"),
    y = max_stops * 0.90,
    label = "Dexter Reed fatal\nshooting incident",
    size = 2
  ) +

  geom_vline(
    xintercept = as.Date("2024-06-14"),
    linetype = "dashed",
    color = "red"
  ) +

  annotate(
    "label",
    x = as.Date("2024-06-14"),
    y = max_stops * 0.98,
    label = "Snelling reports decline\nin traffic stops",
    size = 2,
    hjust = 0.5
  ) +

  geom_vline(
    xintercept = as.Date("2024-08-21"),
    linetype = "dashed",
    color = "red"
  ) +

  annotate(
    "label",
    x = as.Date("2024-08-21"),
    y = max_stops * 0.75,
    label = "Injustice Watch investigation/Bolts:\nNearly 200,000 undocumented arrests",
    size = 2,
    hjust = 0.5
  ) +

  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b"
  ) +

  scale_y_continuous(
    labels = scales::comma
  ) +

  labs(
    title = "Illinois law enforcement traffic stops by month, 2024\n(excluding the Chicago Police Department)",
    x = NULL,
    y = "Number of stops",
    caption = "Source: FOIA Request Results • Graphic by Tiasia Saunders"
  ) +

  theme_minimal() +

  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    ),
    plot.title = element_text(
      face = "bold"
    ),
    plot.caption = element_text(
      hjust = 1,
      size = 8,
      color = "gray40"
    )
  )
```

- The decline in June could be related to the 200,000 undocumented
  traffic stops and on this day Chicago Police Superintendent announced
  at a press release the decrease in traffic stops.

```{r}
### need to create the vehicle make table, will do this tomorrow 
```

#### Per Capital Calculations to get a true reflection of the data, I need to account for population size when looking at the number of stops by race.

```{r}
##  REDO THIS CHUNK calculating for the per capital of total stops by race 
 census_race <- tibble(
  race = c(
    "White",
    "Black",
    "American Indian",
    "Asian",
    "NHPI",
    "Hispanic"
  ),
  population_number = c(
    7472751,
    1775612,
    16561,
    747280,
    2959,
    2337410
  )
)

per_capita_CPD_race <- traffic_CPD_24 |>
  count(race, name = "total_stops") |>
  left_join(census_race, by = "race") |>
  mutate(
    stops_per_100k = round((total_stops / population_number) * 100000, 1)
  ) |>
  arrange(desc(stops_per_100k))

per_capita_CPD_race

```

```{r}
census_race <- census_race |>
  mutate(
    race = recode(race,
      "Native Hawaiian or Other Pacific Islander" = "NHPI"
    )
  )

traffic_24 <- traffic_24 |> 
  mutate(
    driver_race = recode(driver_race, 
      "Native Hawaiian or Other Pacific Islander" = "NHPI"
    )
  )
```

```{r}
unique(traffic_24$race)
unique(census_race$race)
```

```{r}
## making graph to highlight the per capita rate for CPD traffic stops 
 per_capita_CPD_race <- per_capita_CPD_race |>
  mutate(
    highlight = if_else(
      race %in% c("Black", "Hispanic", "White", "Asian"),
      "Major group",
      "Small population group"
    )
  )

ggplot(
  per_capita_CPD_race,
  aes(
    x = reorder(race, stops_per_100k),
    y = stops_per_100k,
    fill = highlight
  )
) +
  geom_col() +

  geom_text(
    aes(label = comma(round(stops_per_100k))),
    hjust = -0.1,
    size = 3
  ) +

  coord_flip() +

  scale_fill_manual(
    values = c(
      "Major group" = "#B22222",
      "Small population group" = "gray70"
    )
  ) +

  scale_y_continuous(
    labels = comma,
    expand = expansion(mult = c(0, .15))
  ) +

  labs(
    title = "Black residents experienced the highest traffic stop rate by the\nChicago Police Department among Illinois' largest racial groups",
    subtitle = "Traffic stops per 100,000 residents, 2024",
    x = NULL,
    y = "Stops per 100,000 residents",
    caption = "Source: Illinois traffic stop data and 2020 Census population estimates.\n Native Hispanics or other Pacific Islander (NHPI) and American Indian (AIAN) rates are based on substantially smaller population counts."
  ) +

  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
  plot.caption = element_text(size = 4, color = "gray40", hjust = 1,   margin = margin(t = 5, r = 1, b = 0, l = 0)), 
      
  )

```

- **THIS HEADLINE IS CONFUSING BECAUSE THE BLACK RESIDENTS ARE THIRD ON
  THE CHART AND YET IT SAYS THEY ARE THE LARGEST – EVEN THOUGH YOU'RE
  QUALIFYING IF BY LARGEST RACIAL GROUP, THE WORDING CLASHES WITH THE
  CHART, SO REVISE**
- CPD traffic stop rates were highest among **STOP USING THESE ACRONYMS.
  SPELL OUT Native Hispanics or other Pacific Islander NOT NHPI**
  ~~NHPI~~ and American Indian residents, though those rates are based
  on much smaller population counts. Among larger racial groups, Black
  residents had the highest stop rate, with about 7,374 stops per
  100,000 residents.
  - ~~Should I take those out or just keep the grey to highlight the
    point of the graph~~?

```{r}
## doing per capital for the other law enforcement agencies 
per_capita_traffic_race <- traffic_24 |>
  count(driver_race, name = "total_stops") |>
  left_join(census_race, by = c("driver_race" = "race")) |>
  mutate(
    stops_per_100k = round((total_stops / population_number) * 100000, 1)
  ) |>
  arrange(desc(stops_per_100k))

per_capita_traffic_race

```

```{r}
## making graph for that 
per_capita_traffic_race <- per_capita_traffic_race |>
  mutate(
    highlight = if_else(
      driver_race %in% c("Black", "Hispanic", "White", "Asian"),
      "Major group",
      "Small population group"
    )
  )

ggplot(
  per_capita_traffic_race,
  aes(
    x = reorder(driver_race, stops_per_100k),
    y = stops_per_100k,
    fill = highlight
  )
) +
  geom_col() +
  geom_text(
    aes(label = comma(round(stops_per_100k))),
    hjust = -0.1,
    size = 3
  ) +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "Major group" = "#B22222",
      "Small population group" = "gray70"
    )
  ) +
  scale_y_continuous(
    labels = comma,
    expand = expansion(mult = c(0, .15))
  ) +
  labs(
    title = "Black residents had the highest stop rate among Illinois' largest racial groups",
    subtitle = "Traffic stops per 100,000 residents, 2024",
    x = NULL,
    y = "Stops per 100,000 residents",
    caption = "Source: Illinois traffic stop data and 2020 Census population estimates.\nNHPI and AIAN rates are based on substantially smaller population counts."
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(
      size = 6,
      color = "gray40",
      hjust = 1,
      margin = margin(t = 5, r = 1, b = 0, l = 0)
    )
  )


## go back ad fix subtitle figure out way to point out NHPI + American Indian, or suggest taking it out 
```

- Illinois state traffic stop rates were highest among NHPI and American
  Indian residents, though those rates are based on much smaller
  population counts. Among larger racial groups, Black residents had the
  highest stop rate, with about 7,374 stops per 100,000 residents,
  compared with 580 per 100,000 white residents.
  - Should I take those out or just keep the grey to highlight the point
    of the graph?

##### Veil of Darkness Element (Very Basic Level)

```{r}
## categorizing my time frames in my 2024 CPD traffic + other traffic datasets 

traffic_CPD_24 <- traffic_CPD_24 |>
  mutate(
    stop_hour = hour(timestop),
    light_period = case_when(
      stop_hour >= 6 & stop_hour < 17 ~ "Daytime",
      stop_hour >= 22 | stop_hour < 4 ~ "Nighttime",
      TRUE ~ "Excluded transition hours"
    )
  )

traffic_CPD_24 |>
  count(light_period)

traffic_CPD_24 <- traffic_CPD_24 |>
  mutate(
    race = recode(
      race,
      "American Indian" = "AIAN"
    )
  )

veil_race_CPD_24 <- traffic_CPD_24 |>
  filter(light_period != "Excluded transition hours") |>
  count(light_period, race) |>
  group_by(light_period) |>
  mutate(percent = n / sum(n) * 100) |>
  ungroup()

veil_race_CPD_24
```

```{r}

veil_race_CPD_24 <- traffic_CPD_24 |>
  mutate(
    stop_hour = hour(timestop),
    light_period = case_when(
      stop_hour >= 6 & stop_hour < 17 ~ "☀ Daytime",
      stop_hour >= 22 | stop_hour < 4 ~ "☾ Nighttime",
      TRUE ~ "Excluded transition hours"
    )
  ) |>
  filter(light_period != "Excluded transition hours") |>
  count(light_period, race) |>
  group_by(light_period) |>
  mutate(
    percent = n / sum(n) * 100,
    race_ordered = reorder_within(race, percent, light_period),
    highlight = if_else(race %in% c("Black", "Hispanic"), "Highlighted", "Other")
  ) |>
  ungroup()

ggplot(
  veil_race_CPD_24,
  aes(
    x = race_ordered,
    y = percent,
    fill = highlight
  )
) +
  geom_col() +
  geom_text(
    aes(label = paste0(round(percent, 1), "%")),
    hjust = -0.1,
    size = 3
  ) +
  coord_flip() +
  facet_wrap(~ light_period, scales = "free_y") +
  scale_x_reordered() +
  scale_fill_manual(
    values = c(
      "Highlighted" = "#B22222",
      "Other" = "gray75"
    )
  ) +
  scale_y_continuous(
    labels = label_percent(scale = 1),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Racial breakdown of CPD traffic stops during\n daylight and nighttime hours",
    subtitle = "Daytime includes 6 a.m.– 5 p.m.; nighttime includes 10 p.m.– 4 a.m.",
    x = NULL,
    y = "Share of stops",
    caption = "Source: Illinois traffic stop data • Graphic by Tiasia Saunders"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold", size = 12),
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 7, color = "gray40", hjust = 1),
    panel.grid.major.y = element_blank()
  )
```

- **DOESNT THAT CONTRADICT YOUR PREVIOUS CHART?** Black drivers
  accounted for the largest share of Chicago traffic stops during both
  daytime (48.0%) and nighttime (41.0%) hours.
- Hispanic drivers' share increased from 31.7% during the day to 37.0%
  at night.
- White drivers accounted for 16.2% of daytime stops and 13.0% of
  nighttime stops.
- At first I was a little confused, but then I realized the racial
  distribution of stops in Chicago differs substantially from statewide
  Illinois traffic-stop patterns, where White drivers make up the
  majority of stops.

```{r}
ggsave(
  "traffic_stops.png",
  width = 8,
  height = 6,
  dpi = 300
)
```

```{r}
## calculating the daytime/nighttime for other traffic data 

traffic_24 <- traffic_24 |>
  mutate(
    stop_hour = as.numeric(substr(time_of_stop, 1, 2)),
    light_period = case_when(
      stop_hour >= 6 & stop_hour < 17 ~ "Daytime",
      stop_hour >= 22 | stop_hour < 4 ~ "Nighttime",
      TRUE ~ "Excluded transition hours"
    )
  )

traffic_24 <- traffic_24 |>
  mutate(
    driver_race = recode(
      driver_race,
      "American Indian" = "AIAN"
    )
  )

veil_race_traffic_24 <- traffic_24 |>
  filter(light_period != "Excluded transition hours") |>
  count(light_period, driver_race) |>
  group_by(light_period) |>
  mutate(percent = n / sum(n) * 100) |>
  ungroup()

veil_race_traffic_24
```

```{r}
## doing the same workflow for the general Illinois traffic stop data 
traffic_24 <- traffic_24 |>
  mutate(
    stop_hour = as.numeric(substr(time_of_stop, 1, 2)),
    light_period = case_when(
      stop_hour >= 6 & stop_hour < 17 ~ " ☀ Daytime",
      stop_hour >= 22 | stop_hour < 4 ~ " ☾ Nighttime",
      TRUE ~ "Excluded transition hours"
    )
  )
```

```{r}
## making the graphic for that 

veil_race <- traffic_24 |>
  filter(light_period != "Excluded transition hours") |>
  count(light_period, driver_race) |>
  group_by(light_period) |>
  mutate(
    percent = n / sum(n) * 100,
    race_ordered = reorder_within(driver_race, percent, light_period),
    highlight = if_else(driver_race %in% c("Black", "Hispanic"), "Highlight", "Other")
  ) |>
  ungroup()

ggplot(veil_race, aes(x = race_ordered, y = percent, fill = highlight)) +
  geom_col() +
  geom_text(
    aes(label = paste0(round(percent, 1), "%")),
    hjust = -0.1,
    size = 3
  ) +
  coord_flip() +
  facet_wrap(~ light_period, scales = "free_y") +
  scale_x_reordered() +
  scale_fill_manual(
    values = c(
      "Highlight" = "#B22222",
      "Other" = "gray75"
    )
  ) +
  scale_y_continuous(
    labels = label_percent(scale = 1),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Racial breakdown of traffic stops during daylight and nighttime hours",
    subtitle = "Daytime: 6 a.m.–5 p.m.; Nighttime: 10 p.m.–4 a.m.",
    x = NULL,
    y = "Share of stops",
    caption = "Source: Illinois traffic stop data • Graphic by Tiasia Saunders"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold", size = 12),
    plot.title = element_text(face = "bold"),
    plot.caption = element_text(size = 7, color = "gray40", hjust = 1),
    panel.grid.major.y = element_blank()
  )
```

- **label this as the illinois police departments data not the chicago
  police**
- During the day, Black drivers account for 18.8 percent of stops and at
  night they account for 27.1 percent of stops
  - That's an increase of 8.3 percentage points, which is a 44 percent
    increase in the share of stops that Black drivers account for at
    night compared to the day
- On the other hand, white drivers fall from 58.1 percent of stops
  during the day to 46.8 percent at night
  - That's a decrease of 11.3 percentage points, which is a 19.5 percent
    decrease in the share of stops that white drivers account for at
    night compared to the day

### Conclusion (thus far):

- CPD Traffic stops by racial disparities breakdown:
  - Black drivers represented the largest share of traffic stops during
    both daytime and nighttime hours, accounting for 48.0% of stops
    during the day and 41.0% at night, while Hispanic drivers' share
    increased from 31.7% during the day to 37.0% at night, and White
    drivers accounted for 16.2% of daytime stops and 13.0% of nighttime
    stops.
- Illinois Traffic stops by other law enforcement agencies:
  - Black drivers represented a larger share of traffic stops during
    nighttime hours than during daytime hours, increasing from 18.8% of
    stops to 27.1%, while White drivers' share declined from 58.1% to
    46.8%.
- Per capital calculations:
  - CPD traffic stop rates were highest among NHPI and American Indian
    residents, though those rates are based on much smaller population
    counts (should I take them out). Among larger racial groups, Black
    residents had the highest stop rate, with about 7,374 stops per
    100,000 residents, compared with 580 per 100,000 white residents.
  - Illinois state traffic stop rates were highest among NHPI and
    American Indian residents, though those rates are based on much
    smaller population counts. Among larger racial groups, Black
    residents had the highest stop rate, with about 7,374 stops per
    100,000 residents, compared with 580 per 100,000 white residents.
  - As for the general Illinois law enforcement agencies data, Among
    larger racial groups, Black residents had the highest stop rate,
    with about 7,374 stops per 100,000 residents, compared with 580 per
    100,000 white residents.
- Overall, I would say the veil of darkness element does seem to be
  supported by the data, as there is a noticeable increase in the share
  of stops that Black drivers account for at night compared to the day,
  and a noticeable decrease in the share of stops that white drivers
  account for at night compared to the day. However, I would say that
  this is not a definitive conclusion, as there could be other factors
  at play that are contributing to these trends, such as differences in
  driving patterns or law enforcement practices during different times
  of the day.
- If I had the opportunity to expand on this project, I would definitely
  to strengthen the veil of darkness angle.
- Overall, I would say that the data does seem to support the idea that
  there are racial disparities in traffic stops, and that these
  disparities are more pronounced during nighttime hours, which is
  consistent with the veil of darkness theory. However, I would also say
  that there are likely other factors at play that are contributing to
  these trends, and that further research would be needed to fully
  understand the underlying causes of these disparities.

### How I could expand on this:

- I could create a more specific formula to account for actual sunset
  and sunrise times throughout the year by connecting weather app data,
  this would be more specific to the veil of darkness element and would
  be more accurate than just categorizing it by hours, but I think for
  the sake of this project and the time frame I have, this is sufficient
  enough to get a general idea of the trends. Should I attempt to do
  this?
- I could also look into the hit rate of drug searches by race and see
  if there are any disparities there, and then also look into the reason
  for stop and see if there are any racial disparities in the reason for
  stop and then also look at the hit rate of drug searches by reason for
  stop and see if there are any disparities there as well.

#### Next Steps (Before presentation):

- I plan to make an interactive map showcasing the amount of stops at
  different agencies locations and the racial breakdown of those stops,
  and then also make a graphic showing the top 10 most common vehicle
  makes that were stopped by CPD in 2024 and the racial breakdown of
  those stops.
- I will think of other potential graphics, but I think I'm good for the
  most part
- I will carefully look back through data analysis to ensure everything
  is correct before transitioning to webpage
- I plan to save these graphics (after looking back through code to make
  sure its correct) and then start to build my webpage for presentation
- I will be using Phoenix Code to create a html, css, and javascript
  webpage for presentation on Saturday
  - I want my end project to look really good so that I can add to
    portfolio and also so that I can share the results with the IPORAE
    members.

##### Business Company Context and Financial Information:

- Tempus AI is one of Chicago's most influential technology companies
  and a key player in the growing field of artificial
  intelligence-driven healthcare. Founded in Chicago, the company uses
  AI and large-scale clinical and genomic data to help physicians make
  more personalized treatment decisions, particularly in cancer care and
  precision medicine.
- For me as a journalist, Tempus is important to watch because it sits
  at the center of several major reporting areas, including AI
  innovation, healthcare, biotechnology, data privacy, and economic
  development. As Chicago continues to emerge as a hub for health
  technology and artificial intelligence, Tempus offers a window into
  how these industries are shaping both the local economy and the future
  of medicine.
- Additionally, Tempus has been in the news for its rapid growth,
  significant funding rounds, and high-profile partnerships with
  healthcare providers and research institutions. Furthermore, its
  first-quarter 2026 revenue, it reported 348.1 million in revenue,
  which exceeded analyst estimates of 345.4 million and it increased by
  36 percent from the same quarter a year prior. The company has raised
  its full-year 2026 revenue forecast to between 1.59 - 1.60 billion.
