---
title: "Collar County Analysis"
author: "Tiasia Saunders"
date: "2025-07-02"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(googlesheets4)
library(googledrive)
library(janitor)
library(rio)
library(data.table)
library(googledrive)
library(knitr)
library(ggplot2)
library(flourishcharts)
library(scales)
library(dplyr)
library(tidyr)
library(DT) 
library(dplyr)
library(ggrepel)
library(lubridate)
library(readxl)
library(htmlwidgets)
library(plotly)
library(tidytext)
library(suncalc)
library(tidygeocoder)
library(splines)
library(stringr)
library(zipcodeR)
library(lutz)
library(pdftables)
library(hms)
```

```{r}
## bringing in traffic_24 data 
traffic_24 <- fread("data/traffic_24.txt", sep="~") |> 
  clean_names()

traffic_CPD_24 <- read_excel("data/traffic_CPD_24.xlsx") |> 
  clean_names()
```

```{r include=FALSE}
## wasnt able to view the traffic_24 data, so i ran the next three code chunks to identify the problem column and fix it 
char_cols <- names(traffic_24)[sapply(traffic_24, is.character)]

for (col in char_cols) {
  cat("Checking:", col, "\n")
  print(try(nchar(traffic_24[[col]]), silent = TRUE))
}
```

```{r include=FALSE}
traffic_24$beat_location_of_stop <-
  iconv(
    traffic_24$beat_location_of_stop,
    from = "",
    to = "UTF-8",
    sub = ""
  )
```

```{r include=FALSE}
nchar(traffic_24$beat_location_of_stop)
```

```{r}
traffic_24$sex <- dplyr::recode(
  as.character(traffic_24[["driver_sex"]]),
  "1" = "Male",
  "2" = "Female",
  "3" = "Non-Binary",
  "9" = "Unknown"
)

traffic_24$race <- dplyr::recode(
  as.character(traffic_24[["driver_race"]]),
  "1" = "White",
  "2" = "Black",
  "3" = "American Indian",
  "4" = "Hispanic",
  "5" = "Asian",
  "6" = "NHPI"
)

traffic_24$vehicle_search_conducted <- dplyr::recode(
  as.character(traffic_24[["vehicle_search_conducted"]]),
  "1" = "Yes",
  "2" = "No"
)

```

```{r}
traffic_CPD_24 <- traffic_CPD_24 |>
  mutate(
    sex = recode(as.character(drsex),
      "1" = "Male",
      "2" = "Female",
      "3" = "Non-Binary", 
      "9" = "Unknown"
    ),
    race = recode(as.character(drrace),
      "1" = "White",
      "2" = "Black",
      "3" = "American Indian",
      "4" = "Hispanic",
      "5" = "Asian",
      "6" = "NHPI"
    ))

```

```{r}
agency_counts <- traffic_24 %>%
  group_by(agency_name) %>%
  summarise(
    total_stops = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(total_stops))
```

```{r}
head(traffic_24)
```

```{r}
traffic_24 <- traffic_24 %>%
  mutate(
    datetime = as.POSIXct(
      paste(date_of_stop, time_of_stop),
      format = "%Y-%m-%d %H:%M",
      tz = "America/Chicago"
    )
  ) 
 
```

```{r}
traffic_24 <- traffic_24 |> 
  mutate(date = as.Date(date_of_stop))
  #rename(date = date_of_stop)

```

```{r}
date_tbl <- traffic_24 %>%
  distinct(date)
```

```{r}

il_lat <- 39.8283
il_lon <- -89.2460

sun_times <- getSunlightTimes(
  date = date_tbl$date,
  lat = il_lat,
  lon = il_lon,
  tz = "America/Chicago"
)
```

```{r}
traffic_24 <- traffic_24 %>%
  left_join(sun_times, by = "date")
```

```{r}
sun_times <- sun_times %>%
  mutate(
      sunrise_time = format(sunrise, "%H:%M"),
    sunset_time  = format(sunset, "%H:%M"), 
    dawn_time = format(dawn, "%H:%M"), 
    dusk_time = format(dusk, "%H:%M")
  )
```

```{r}
traffic_24 <- traffic_24 %>%
  mutate(
    daylight = datetime >= sunrise & datetime <= sunset,

    light_condition = case_when(
      datetime < dawn ~ "Night",
      datetime >= dawn & datetime < sunrise ~ "Morning Twilight",
      datetime >= sunrise & datetime <= sunset ~ "Daylight",
      datetime > sunset & datetime <= dusk ~ "Evening Twilight",
      TRUE ~ "Night"
    )
  )
```

```{r}
### checking light_conditions 

traffic_24 |> 
  count(light_condition, sort = TRUE)
```
```{r}
traffic_24 <- traffic_24  %>%
  mutate(
    vod = ifelse(datetime >= sunrise.x & datetime <= sunset.x, 1, 0)
  )
```

```{r}
## checking stop times and light conditions 
stop_times_vod <- traffic_24 |> 
  select(date,time_of_stop, light_condition) 
```

```{r}
## checking the range of dates in the dataset 
range(traffic_24$date)
```

```{r}
## verifying the vod results 
table(traffic_24$vod, useNA = "ifany")
```

```{r}
traffic_24 %>%
  filter(is.na(vod)) %>%
  select(date, time_of_stop, datetime, sunrise, sunset)
```

```{r}
## taking out the NA values 
traffic_24 <- traffic_24 %>%
  filter(!is.na(datetime))

```

```{r}
## making sure there is no NA values in the datetime column
sum(is.na(traffic_24$datetime))
```

```{r}
## verifying the vod results 
mean(traffic_24$match, na.rm = TRUE)
```
```{r}
## just getting the daytime and nighttime stops 
traffic_24_vod <- traffic_24 %>%
  filter(light_condition %in% c("Daylight", "Night"))
```


```{r}
traffic_24_vod |> 
  select(time_of_stop, sunrise, sunset, vod, light_condition)
```

```{r}
## grouping by agency 
vod_summary <- traffic_24_vod %>%
  group_by(agency_name, race, vod) %>%
  summarise(stops = n(), .groups = "drop")
```

```{r}
vod_wide <- vod_summary %>%
  tidyr::pivot_wider(
    names_from = vod,
    values_from = stops,
    values_fill = 0
  ) %>%
  mutate(
    day = `1`,
    night = `0`,
    night_share = night / (day + night)
  )
```

```{r}
top5_agencies <- traffic_24_vod %>%
  group_by(agency_name) %>%
  summarise(
    total_stops = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(total_stops)) %>%
  slice_head(n = 5)
```

```{r}
traffic_top5 <- traffic_24_vod %>%
  filter(agency_name %in% top5_agencies$agency_name)
```

```{r}
race_vod_top5 <- traffic_top5 %>%
  group_by(agency_name, race, vod) %>%
  summarise(stops = n(), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from = vod,
    values_from = stops,
    values_fill = 0
  ) %>%
  mutate(
    day = `1`,
    night = `0`,
    total = day + night,
    night_share = night / total,
    day_share = day / total
  )

race_vod_top5
```

```{r}

vod_long <- race_vod_top5 %>%
  select(agency_name, race, day_share, night_share) %>%
  pivot_longer(
    cols = c(day_share, night_share),
    names_to = "time_of_day",
    values_to = "share"
  ) %>%
  mutate(
    time_of_day = recode(
      time_of_day,
      day_share = "Day",
      night_share = "Night"
    )
  )
```

```{r}
ggplot(vod_long,
       aes(x = race,
           y = share,
           fill = time_of_day)) +
  geom_col() +
  facet_wrap(~ agency_name) +
  scale_fill_manual(
    values = c(
      "Day" = "grey70",
      "Night" = "#B22222"
    )
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "The Top 5 Agencies Proportion of Daytime and Nighttime Traffic Stops by Race",
    subtitle = "Traffic stops classified using sunrise and sunset; twilight periods excluded",
    x = "Race",
    y = "Share of Stops",
    fill = "Time of Day"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 6),
    legend.position = "top"
  )
```

[ rewrite the below]
- So looking at the top 5 agencies:  
  - Illinois State Police Black traffic stops decrease from day (34700) to (33312)
  - For ISP white people traffic stops decrease from (93880) to (51063)
  - For ISP hispanic people traffic stops decrease from (17180) to (12193)
  
  - For Bolingbrook police Black traffic stops decrease from (3220) to (3068)
  - for white people traffic stops decrease from (3306) to (2205)
  - For Hispanich people it decreases from (3226) to (2799)
  
  - For Cook County Sheriff Black traffic stops increased from (7152) to (8856)
  - For white people traffic stops it decreased from (10540) to (8204)
  - For hispanic people traffic stops it increased from (8123) to (8400)
  
  - For Schamburg Police Black traffic stops decreased from (1302) to (1222)
  - For white people traffic stops it decreased from (7188) to (3535) 
  - For hispanic people it decreased from (3478) to (2858)
  
  - For Palatine Police Black people stops decreased from (774) to (620)
  - For white people traffic stops it decreased from (8406) to (2719)
  - For hispanic people it decreased from (3376) to (1947)
  
  


### CPD VoD Analysis 

```{r}
## looking at top 5 cities, that I will do further analysis on 
traffic_CPD_24 %>%
  filter(!is.na(city_i)) %>%
  count(city_i, sort = TRUE) %>%
  slice_head(n = 5)
```

```{r}
traffic_CPD_24 <- traffic_CPD_24 |>
  mutate(
    date = as.Date(datestop)
  )

```

```{r}
## using CPD data for stops that happened in chicago for the year 2024, starting with CPD since its just one agency and its easier to work with 
chicago_stops <- traffic_CPD_24 |> 
  filter(city_i == "CHICAGO")
```


```{r}
chicago_stops <- chicago_stops |>
  mutate(
    time = format(timestop, "%H:%M:%S"),
    datetime = as.POSIXct(
      paste(datestop, time),
      format = "%Y-%m-%d %H:%M:%S",
      tz = "America/Chicago"
    )
  )
 
```

```{r}
traffic_CPD_24 <- traffic_CPD_24 |> 
  mutate(date = as.Date(date)) 
  # rename(date = date_of_stop)

```

```{r}

chicago_lat <- 41.8818
chicago_lng <- -87.6231


chi_sun_times <- getSunlightTimes(
  date = date_tbl$date,
  lat = chicago_lat,
  lon = chicago_lng,
  tz = "America/Chicago"
)
```

```{r}
chicago_stops <- chicago_stops %>%
  left_join(chi_sun_times, by = "date")
```

```{r}
chi_sun_times <- chi_sun_times %>%
  mutate(
      sunrise_time = format(sunrise, "%H:%M"),
    sunset_time  = format(sunset, "%H:%M"), 
    dawn_time = format(dawn, "%H:%M"), 
    dusk_time = format(dusk, "%H:%M")
  )
```

```{r}
chicago_stops <- chicago_stops %>%
  mutate(
    daylight = datetime >= sunrise & datetime <= sunset,

    light_condition = case_when(
      datetime < dawn ~ "Night",
      datetime >= dawn & datetime < sunrise ~ "Morning Twilight",
      datetime >= sunrise & datetime <= sunset ~ "Daylight",
      datetime > sunset & datetime <= dusk ~ "Evening Twilight",
      TRUE ~ "Night"
    )
  )
```

```{r}
chicago_stops <- chicago_stops  %>%
  mutate(
    vod = ifelse(datetime >= sunrise & datetime <= sunset, 1, 0)
  )
```

```{r}
### checking light_conditions 
chicago_stops |> 
  count(light_condition, sort = TRUE)
```

```{r}
## checking stop times and light conditions 
chi_stop_times_vod <- chicago_stops |> 
  select(date,time, light_condition) 
```

```{r}
## checking the range of dates in the dataset 
range(chicago_stops$date)
```

```{r}
## verifying the vod results 
table(chicago_stops$vod, useNA = "ifany")
```


```{r}
## just getting the daytime and nighttime stops 
chicago_stops_vod <- chicago_stops %>%
  filter(light_condition %in% c("Daylight", "Night"))
```


```{r}
chicago_stops_vod |> 
  select(time, sunrise, sunset, vod, light_condition)
```

```{r}
## grouping by agency 
chi_vod_summary <- chicago_stops_vod %>%
  group_by(race, vod) %>%
  summarise(stops = n(), .groups = "drop")
```

```{r}
chi_vod_wide <- chi_vod_summary %>%
  tidyr::pivot_wider(
    names_from = vod,
    values_from = stops,
    values_fill = 0
  ) %>%
  mutate(
    day = `1`,
    night = `0`,
    night_share = night / (day + night), 
    day_share = day / (day + night) 
  )

chi_vod_wide
```


```{r}

chi_vod_long <- chi_vod_wide %>%
  select(race, day_share, night_share) %>%
  pivot_longer(
    cols = c(day_share, night_share),
    names_to = "time_of_day",
    values_to = "share"
  ) %>%
  mutate(
    time_of_day = recode(
      time_of_day,
      day_share = "Day",
      night_share = "Night"
    )
  )
```

```{r}
ggplot(chi_vod_long,
       aes(x = race,
           y = share,
           fill = time_of_day)) +
  geom_col() +
  scale_fill_manual(
    values = c(
      "Night" = "grey70",
      "Day" = "#B22222"
    )
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    title = "CPD Chicago Proportion of Daytime and Nighttime Traffic Stops by Race",
    subtitle = "Traffic stops classified using sunrise and sunset; twilight periods excluded",
    x = "Race",
    y = "Share of Stops",
    fill = "Time of Day"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 6),
    legend.position = "top"
  )
```





















```{r}
## verifying the vod results 
chicago_stops <- chicago_stops %>%
  mutate(
    match = ifelse(is.na(vod) | is.na(vod_check), NA, vod == vod_check)
  )
```

```{r}
## verifying the vod results 
mean(traffic_24$match, na.rm = TRUE)
```

```{r}
chicago_stops |> 
  select(timestop, sunrise.x, sunset.x, vod)
```

```{r}
chi_vod_summary <- chicago_stops %>%
  group_by(race, vod) %>%
  summarise(stops = n(), .groups = "drop")
```

```{r}
chi_vod_wide <- chi_vod_summary %>%
  tidyr::pivot_wider(
    names_from = vod,
    values_from = stops,
    values_fill = 0
  ) %>%
  mutate(
    day = `1`,
    night = `0`,
    night_share = night / (day + night)
  )
```

```{r}
chi_top5_race <- chicago_stops %>%
  group_by(race) %>%
  summarise(
    total_stops = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(total_stops)) %>%
  slice_head(n = 5)
```


```{r}
chi_race_vod_top5 <- chicago_stops %>%
  group_by(race, vod) %>%
  summarise(stops = n(), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from = vod,
    values_from = stops,
    values_fill = 0
  ) %>%
  mutate(
    day = `1`,
    night = `0`,
    total = day + night,
    night_share = night / total,
    day_share = day / total
  )
```


```{r}
ggplot(chi_race_vod_top5, aes(x = race)) +
  geom_col(aes(y = day), fill = "grey70") +
  geom_col(aes(y = night), fill = "#B22222") +
  coord_flip() +
  labs(
    title = "Chicago Traffic Stops by Race: Day vs. Night",
    x = "Race",
    y = "Number of Stops"
  )

```


